### 2016-06-26 motivation.tex
Secs. 4.1 and 4.2 done.
edits to secs. 1 - 4

\begin{figure}[!h]
\centering
\subcaptionbox {
  {\sc rc} Execution 1
  \label{fig:motiv-eg-1-a}
} [
  0.55\columnwidth
] {
  \includegraphics[scale=0.5]{Figures/motiv-eg-1-a}
}
%\hspace*{0.5in}
\subcaptionbox {
  {\sc rc} Execution 2
  \label{fig:motiv-eg-1-b}
}{
  \includegraphics[scale=0.5]{Figures/motiv-eg-1-b}
}
\caption{A possible execution of the program shown in Fig.~\ref{fig:motiv-eg-1} under
  a read committed isolation level.}
\label{fig:rc-executions}
\end{figure}
The figure depicts an execution as a series of read, write, and commit
operations.  The subscript of an operation indicates the transaction
that executes it\footnote{For clarity, the effects of different
  transactions are shown in different colored backgrounds.} In the
first execution, transaction {\bf Wd1} reads the current balance
(\C{k}) and writes the new balance (\C{k-a1}), but before it commits
transaction {\bf Wd2} executes and commits, writing the new balance
(\C{k-a2}). RC isolation prevent {\bf Wd2} from witnessing the
uncommitted writes of transaction {\bf Wd1}.  Subsequently committing
{\bf Wd1} leads to the loss of {\bf Wd2}'s updates (the so called
\emph{lost update anomaly}~\cite{berenson}), resulting in an incorrect
balance of \C{k-a1}. The second execution describes a similar scenario
with {\bf Wd1} and {\bf Wd2} exchanging their roles.  Clearly, {\sc
  rc} is an excessively weak isolation level for this program because
it loses the updates of one transaction, resulting in the violation of
the post-condition.  A stronger isolation level that prevents lost
updates is required.  \iso{Snapshot Isolation} ({\sc
  si})~\cite{berenson} fits this requirement; {\sc si} effectively
serializes transactions that update a shared data object by aborting
and re-executing a transaction if write-write conflicts are detected
during its commit.\footnote{{\sc si} is not functionally the same as
  {\sc ser}, however; the differences in the two manifest in the \emph{absence}
  of any write-write conflicts.}   Since {\sc si},
unlike {\sc ser}, does not rely on expensive lock-based
concurrency control, it is also more efficient, making it appropriate
appropriate for both {\bf Wd} transactions.

@@

Thinking in terms of anomalies, as described above, is how database
programmers are often encouraged to reason about weak isolation.
Unfortunately, such reasoning does not rest on any sound foundation,
and is thus highly error-prone. Reasoning in terms of how weak
isolation variants are implementated is no better since it requires
application programmers to understand low-level implementation details
of the database that are naturally far removed from the application
semantics. An attractive (and principled) alternative in this context
is a formal proof system that combines declarative reasoning about
isolation guarantees with operational reasoning about programs. We
demonstrate how our proof system makes this possible in the context of
the current example.

@@

First, we note that the example in Fig.~\ref{fig:motiv-eg-1} is a
concurrent program, and hence is amenable to \emph{rely-guarantee}
style reasoning~\cite{rgjones}, a compositional proof technique that
allows us to reason about one thread at a time by abstracting away
interference induced by other threads (collectively called \emph{the
  environment}) into a \emph{rely} relation. In an ordinary concurrent
program, every environment step is a valid interference in the current
thread. However, in the presence of transactions executing under weak
isolation, determining what constitutes an interference is a
non-trivial problem. For example, inside a serializable transaction
there is no interference, whereas within a transaction executing under
\iso{Read Uncommitted}, the weakest isolation level, all interferences
are valid. Between these two extremes are various levels of isolation
that admit some interferences while prohibiting others.  For example,
\iso{Read Committed} isolation admits interference of a committed
transaction, but not that of an uncommitted transaction.
\iso{Snapshot Isolation} admits interference from committed
non-conflicting transactions, and so on.

For the rely-guarantee approach to be useful, it has to thus constrain
allowed interference in accordance with the chosen level of isolation
associated with a transaction. The first key observation we make is
that we can constrain interference by constraining the definition of
the \emph{happens-before} ($\hbZ$) relation between transactions, or
their respective constituents.  To do so, we axiomatize the $\hbZ$
relation to capture the interference characteristics of different
isolation levels as their specifications ($\psi$). For example, a
\iso{Read Committed} specification allows an operation $\eta_1$ of
$T_1$ to happen before an operation $\eta_2$ of $T_2$ only if every
operation in $T_1$ (including its commit) happens before $\eta_2$ (we
let $T_1 \hboar \eta_2$ denote this). The {\sc rc} specification
does not require $T_1$ to execute and commit before all operations of
$T_2$, thus allowing it to interfere in $T_2$. The \iso{Snapshot
  Isolation} specification requires two transactions that write to the
same variable to be related by happens-before.  This effectively
prohibits interference due to actions in $T_1$ being interleaved in
$T_2$ (or, vise versa).  We formalize these properties in Fig.~\ref{fig:interference-ex}\footnote{Note that specifications
    presented here are solely for illustration. The actual
    specifications described in \S\ref{sec:ansi-isolation} are more nuanced.}.

@@

The specification of an isolation level encodes its interference
characteristic as constraints over the $\hbZ$ relation. However, for this
to be useful in reasoning about programs, the rely-guarantee framework
should be able to use the specification to determine if an
interference is valid or invalid, allowing the programmer to only
focus on the former. Our second key observation is that this is
possible if the reasoning framework adequately tracks $\hbZ$ at each
program point, while preserving $\hbZ$ constraints as invariants
between program points. An interference that leads to the violation of
$\hbZ$ constraints (i.e., an invalid interference) is thus
automatically prohibited. For instance, consider the program point
after the write to \C{C} in {\bf Wd1}. The expected invariant ($\phi$) at
that program point is shown below ($\committed$ stands for
``committed''):

\begin{smathpar}
\begin{array}{l}
  \neg\committed(\C{Wd1}) \conj (\neg\committed(\C{Wd2}) \Rightarrow
  \C{C = k-a1}) 
                \\
      \C{Wd1} \wrstoar \C{C} \conj (\committed(\C{Wd2})
                \Rightarrow \C{C = k-a1-a2})\}
\end{array}
\end{smathpar}

@@

We have thus far assumed a strongly consistent ({\sc SC}) store that,
in the absence of transactions with special isolation requirements,
makes the effects of any operation immediately visible to all
subsequent operations. The natural behavior of {\sc SC} to totally
order all operations w.r.t the $\hbZ$ relation could be in conflict
with the constraints imposed by weak isolation. To meet the
requirements of weaker isolation levels, say {\sc rc}, the store has
to adapt itself to \emph{hide} the effects of concurrent transactions
until they are committed. Once committed though, effects need to be
immediately and perennially visible to subsequent operations. This
semantics can be built into the reasoning framework, leading to a proof
system tailor-made for such stores. However, for the reasoning
framework to be truly useful, it has to be parametric over different
store consistency semantics, supporting stores that are weaker than
{\sc sc} (e.g., eventually consistent), and should be able to
reconcile conflicts between consistency and isolation
constraints.  We demonstrate how our reasoning framework
makes this possible in the following sections.

>>>>

\begin{figure}[!h]
\centering
\subcaptionbox {
  {\sc rc} Execution 1
  \label{fig:motiv-eg-1-a}
} [
  0.55\columnwidth
] {
  \includegraphics[scale=0.5]{Figures/motiv-eg-1-a}
}
%\hspace*{0.5in}
\subcaptionbox {
  {\sc rc} Execution 2
  \label{fig:motiv-eg-1-b}
}{
  \includegraphics[scale=0.5]{Figures/motiv-eg-1-b}
}
\caption{A possible execution of the program shown in Fig.~\ref{fig:motiv-eg-1} under
  a read committed isolation level.}
\label{fig:rc-executions}
\end{figure}
The figure depicts an execution as a series of read, write, and commit
operations.  The subscript of an operation indicates the transaction
that executes it.\footnote{For clarity, the effects of different
  transactions are shown in different colored backgrounds.} In the
first execution, transaction {\bf Wd1} reads the current balance
(\C{k}) and writes the new balance (\C{k-a1}), but before it commits
transaction {\bf Wd2} executes and commits, writing the new balance
(\C{k-a2}). RC isolation prevent {\bf Wd2} from witnessing the
uncommitted writes of transaction {\bf Wd1}.  Subsequently committing
{\bf Wd1} leads to the loss of {\bf Wd2}'s updates (the so called
\emph{lost update anomaly}~\cite{berenson}), resulting in an incorrect
balance of \C{k-a1}. The second execution describes a similar scenario
with {\bf Wd1} and {\bf Wd2} exchanging their roles.  Clearly, {\sc
  rc} is an excessively weak isolation level for this program because
it loses the updates of one transaction, resulting in the violation of
the post-condition.  A stronger isolation level that prevents lost
updates is required.  \iso{Snapshot Isolation} ({\sc
  si})~\cite{berenson} fits this requirement; {\sc si} effectively
serializes transactions that update a shared data object by aborting
and re-executing a transaction if write-write conflicts are detected
during its commit.\footnote{{\sc si} is not functionally the same as
  {\sc ser}, however; the differences in the two manifest in the \emph{absence}
  of any write-write conflicts.}   Since {\sc si},
unlike {\sc ser}, does not rely on expensive lock-based
concurrency control, it is also more efficient, making it appropriate
for both {\bf Wd} transactions.

@@

Thinking in terms of anomalies, as described above, is how database
programmers are often encouraged to reason about weak isolation.
Unfortunately, such reasoning does not rest on any sound foundation,
and is thus highly error-prone. Reasoning in terms of how weak
isolation variants are implementated is no better since it requires
programmers to understand low-level implementation details
of the database that are naturally far removed from the application
semantics. An attractive (and principled) alternative in this context
is a formal proof system that combines declarative reasoning about
isolation guarantees with operational reasoning about programs. We
demonstrate how our proof system makes this possible in the context of
the current example.

@@

First, we note that the example in Fig.~\ref{fig:motiv-eg-1} is a
concurrent program, and hence is amenable to \emph{rely-guarantee}
style reasoning~\cite{rgjones}, a compositional proof technique that
allows us to reason about the behaviour of individual threads by
abstracting away interference induced by other threads (collectively
called \emph{the environment}) into a \emph{rely} relation. In an
ordinary concurrent program, every environment step is a valid
interference in the current thread. However, in the presence of
transactions executing under weak isolation, determining what
constitutes an interference is a non-trivial problem. For example,
inside a serializable transaction there is no interference, whereas
within a transaction executing under a \iso{Read Uncommitted}
isolation level, the weakest isolation level, all interferences are
valid. Between these two extremes are various levels of isolation that
admit some interferences while prohibiting others.  For example,
\iso{Read Committed} isolation admits interference of a committed
transaction, but not that of an uncommitted transaction.
\iso{Snapshot Isolation} admits interference from committed
non-conflicting transactions, and so on.

For the rely-guarantee approach to be useful, it has to thus constrain
allowed interference in accordance with the chosen level of isolation
associated with a transaction. The first observation we make is
that we can constrain interference by constraining the definition of
the \emph{happens-before} ($\hbZ$) relation that dictates visibility
actions between transactions, or their respective constituents.  To do
so, we axiomatize the $\hbZ$ relation to capture the interference
characteristics of different isolation levels as their specifications
($\psi$). For example, a \iso{Read Committed} specification allows an
operation $\eta_1$ of $T_1$ to happen before an operation $\eta_2$ of
$T_2$ only if every operation in $T_1$ (including its commit) happens
before $\eta_2$ (we let $T_1 \hboar \eta_2$ denote this). The 
specification does not require $T_1$ to execute and commit before all
operations of $T_2$, thus allowing it to interfere in $T_2$. The
\iso{Snapshot Isolation} specification requires two transactions that
write to the same variable to be related by happens-before.  This
effectively prohibits interference due to actions in $T_1$ being
interleaved in $T_2$ (or, vise versa).  We formalize these properties
in Fig.~\ref{fig:interference-ex}.\footnote{Note that specifications
  presented here are solely for illustration. The actual
  specifications described in \S\ref{sec:ansi-isolation} are more
  nuanced.}.

@@

The specification of an isolation level encodes its interference
characteristic as constraints over the $\hbZ$ relation. However, for
this to be useful in reasoning about programs, the rely-guarantee
framework should be able to use the specification to determine if an
interference is valid or invalid, allowing the programmer to only
focus on the former. Our second observation is that this is possible
if the reasoning framework adequately tracks $\hbZ$ at each program
point, while preserving $\hbZ$ constraints as invariants between
program points. An interference that leads to the violation of $\hbZ$
constraints (i.e., an invalid interference) is thus automatically
prohibited. For instance, consider the program point after the write
to \C{C} in {\bf Wd1}. The expected invariant ($\phi$) at that program
point is shown below ($\committed$ stands for ``committed''):

\begin{smathpar}
\begin{array}{l}
  \neg\committed(\C{Wd1}) \conj (\neg\committed(\C{Wd2}) \Rightarrow
  \C{C = k-a1}) 
                \\
      \C{Wd1} \wrstoar \C{C} \conj (\committed(\C{Wd2})
                \Rightarrow \C{C = k-a1-a2})
\end{array}
\end{smathpar}

@@

We have thus far assumed a strongly consistent ({\sc sc}) store that,
in the absence of transactions with special isolation requirements,
makes the effects of any operation immediately visible to all
subsequent operations. The natural behavior of {\sc sc} to totally
order all operations w.r.t the $\hbZ$ relation could be in conflict
with the constraints imposed by weak isolation. To meet the
requirements of weaker isolation levels, say {\sc rc}, the store has
to adapt itself to \emph{hide} the effects of concurrent transactions
until they are committed. Once committed though, effects need to be
immediately visible to subsequent operations. This semantics can be
built into the reasoning framework, leading to a proof system
tailor-made for such stores. However, for the reasoning framework to
be truly useful, it has to be parametric over different store
consistency semantics, supporting stores that are weaker than {\sc sc}
(e.g., eventually consistent), and should be able to reconcile
conflicts between consistency and isolation constraints.  We
demonstrate how our reasoning framework makes this possible in the
following sections.

### 2016-06-26 opsem.tex
edits to secs. 1 - 4

\section{\txnimp}
\label{sec:syntax}

\subsection{\txnimp: Syntax and Semantics}
\label{sec:opsem}

@@

Fundamental to our development is the notion of a trace invariant
($\I$). $\I$ is a function from traces ($\E$) to propositions
($\texttt{Prop}$) that define well-formedness constraints over traces.
The machine takes a step only if the resulting trace satisfies the
constraints imposed by $\I$. This behaviour is captured by the
auxiliary reduction rule \rulelabel{E-Aux} that factors out the trace
extension aspect of the evaluation by abstracting away the
operation-specific behaviour as a function that generates an
appropriate effect. We let $\mathcal{F}$ denote this function.
\rulelabel{E-Aux} uses $\mathcal{F}$ to generate a new effect and
extend the trace ($\E = (\A,\visZ)$) \emph{only if} the
well-formedness constraints imposed by $\I$ on $\E$ (i.e., $\I(\E')$)
are satisfied. Otherwise, it gets stuck. In an execution that runs to
completion, every small-step preserves the well-formedness of a trace,
thus ensuring the invariance of $\I$.  Note that the semantics makes
no assumptions about $\I$ other than its type. As such, it can be
instantiated with any trace-parametric proposition that expresses
constraints over the given trace. For instance, consider the
$\psi_{RC}$ specification from \S\ref{sec:motivation}, but with
bounded $T_1$ and $T_2$ instantiated with \C{Wd1} and \C{Wd2},
respectively. The instantiated specification is the following
proposition:

@@

\noindent It is easy to intepret the above proposition in the context of a trace
$\E$ that captures an execution of the program in
Fig.~\ref{fig:motiv-eg-1}. Such a trace-parametric proposition can be
used to instantiate the trace invariant $\I$ in Fig.~\ref{fig:txnimp}.
The resultant operational semantics describes an abstract machine that
gets stuck if an operation of \C{Wd2} is executed in a state that
incorporates some, but not all the effects (including the \C{COMMIT})
of \C{Wd1}. 
% While the machine takes a step only if the constraints are
% satisfied, it neither defines nor explicitly assumes an oracle to
% check satisfaction.

As described in \S\ref{sec:motivation}, the semantics of various
isolation levels can be captured as constraints over the
happens-before ($\hbZ$) relation. $\hbZ$ is however a derived relation
in our model, composed of more fundamental \emph{session order}
($\soZ$) and \emph{visibility} ($\visZ$) relations, which shall
henceforth replace $\hbZ$ in the discussion.  A session order relation
captures the sequential order of operations within a transaction. In
particular, it relates two effects, $\eta_1$ and $\eta_2$, such that
$\txn(\eta_1) = \txn(\eta_2)$ and $\id(\eta_1) < \id(\eta_2)$.  The
semantics assigns monotonically increasing identifiers to effects, as
defined by the $\id(\eta) > \maxId(\A)$ condition of
\rulelabel{E-Aux}.  Evaluation contexts ($\ectx_i$) for
transaction-bound terms are defined so as to enforce a deterministic
sequential order of execution within a transaction, leading to a
deterministic total order among effect ids, which defines the session
order relation. Visibility ($\visZ$) on the other hand relates effects
across concurrent transactions.  Intuitively, $\visZ$ relates $\eta_1$
to $\eta_2$ if and only if $\eta_1$ was \emph{visible} to the
operation that generated $\eta_2$ during its execution, thus effecting
its return value ($\rval(\eta_2)$). For example, a read operation over
\C{X} may pick the value ($\rval$) of the write effect with highest id
among the visible effects (this is made possible by appropriately
defining $\interp{\cdot}$ in \rulelabel{E-Var}, as we show later).
Thus, the value of a read depends on what write effects it can
witness. An operation can only witness the effects of already
concluded operations, which varies between executions due to the
non-deterministic order of evaluating the parallel composition of
transactions. A more notable source of non-determinism, however, is
the \rulelabel{E-Aux} rule, which allows the machine to expose an
arbitrary subset ($S$) of existing effects ($\A$) to the incoming
operation. In other words, the machine is not obligated to reveal the
effects of all previous operations to an incoming operation. This
relaxation allows the abstract machine to model the semantics of
weakly consistent data stores. For instance, operations issued to an
eventually consistent (EC) replicated store could be dispatched to
different replicas whose states may not be in any well-defined
relationship. By allowing operations to witness arbitrary subsets of
the global state, the semantics models the weak visibility properties
of such stores. Stronger visiblity properties needed for stronger
isolation can however be recovered by imposing well-formedness
constraints over $\visZ$ via the trace invariant ($\I$). Since the
abstract machine is obligated to satisfy $\I$ at every step of the
execution, operations are guaranteed to experience the level of
isolation specified by $\I$.  Thus, in executions that run to
completion, the abstract machine models a store that provides the
required levels of isolation.  Notably, the machine achieves this
without defining an operation semantics for isolation levels, solely
by relying on their declarative characterization as trace
well-formedness constraints.  \S\ref{sec:ansi-isolation} shows the
specification of various ANSI SQL isolation levels stated as trace
well-formedness constraints.

@@

As described previously, \rulelabel{E-Aux} abstracts away the
operation-specific behaviour of a machine step as a function ($\F$)
that accepts a set ($S$) of effects chosen by the machine to make
visible to the operation, interprets the operation w.r.t $S$, and
returns an appropriate effect that encodes its return value. Rules
\rulelabel{E-Var}, \rulelabel{E-Asgn} and \rulelabel{E-Commit} define
such functions for read, write and commit operations, respectively.
The effect returned by the function in each case includes its
transaction id ($T_i$) along with an arbitrarily chosen effect id
($j$) that is later verified to be unique in \rulelabel{E-Aux}. The
$\rval$ for a write is the value being written, and for commit it is
$\bot$. In case of a read, the value read depends on the how the read
operation chooses to interpret the given set ($S$) of visible effects.
The interpretation may depend on the application semantics. For
example, a monotonically increasing counter application may choose to
let a write with the largest value determine the value of a
read. A more straightforward interpretation has already been described
before. To accommodate multiple interpretations, the semantics
is made parametric over an interpretation function ($\interp{\cdot}$)
that accepts a set of effects and a variable name, and returns the
value associated with the variable. A straightforward interpretation
function that chooses the last write (i.e., write with largest id) is
shown below:

>>>>

\section{\txnimp: Syntax and Semantics}
\label{sec:opsem}

\label{sec:syntax}

@@

Fundamental to our development is the notion of a trace invariant
($\I$). $\I$ is a function from traces ($\E$) to first-order logical
formula ($\texttt{Prop}$) that define well-formedness
constraints over traces.  The machine takes a step only if the
resulting trace satisfies the constraints imposed by $\I$. This
behaviour is captured by the auxiliary reduction rule
\rulelabel{E-Aux} that factors out the trace extension aspect of the
evaluation by abstracting away the operation-specific behaviour as a
function that generates an appropriate effect. We let $\mathcal{F}$
denote this function.  \rulelabel{E-Aux} uses $\mathcal{F}$ to
generate a new effect and extend the trace ($\E = (\A,\visZ)$)
\emph{only if} the well-formedness constraints imposed by $\I$ on $\E$
(i.e., $\I(\E')$) are satisfied. Otherwise, it gets stuck. In an
execution that runs to completion, every small-step preserves the
well-formedness of a trace, thus ensuring the invariance of $\I$.
Note that the semantics makes no assumptions about $\I$ other than its
type. As such, it can be instantiated with any trace-parametric
proposition that expresses constraints over the given trace. For
instance, consider the $\psi_{RC}$ specification from
\S\ref{sec:motivation}, but with bounded $T_1$ and $T_2$ instantiated
with \C{Wd1} and \C{Wd2}, respectively. The instantiated specification
is the following term:

@@

\noindent It is easy to interpret the above specification in the
context of a trace $\E$ that captures an execution of the program in
Fig.~\ref{fig:motiv-eg-1}. Such a trace-parametric formula can be
used to instantiate the trace invariant $\I$ in Fig.~\ref{fig:txnimp}.
The resultant operational semantics describes an abstract machine that
gets stuck if an operation of \C{Wd2} is executed in a state that
incorporates some, but not all the effects (including the \C{COMMIT})
of \C{Wd1}.
% While the machine takes a step only if the constraints are
% satisfied, it neither defines nor explicitly assumes an oracle to
% check satisfaction.

As described in \S\ref{sec:motivation}, the semantics of various
isolation levels can be captured as constraints over the
happens-before ($\hbZ$) relation. $\hbZ$ is however a derived relation
in our model, composed of more fundamental \emph{session order}
($\soZ$) and \emph{visibility} ($\visZ$) relations, which shall
henceforth replace $\hbZ$ in the discussion.  A session order relation
captures the sequential order of operations within a transaction. In
particular, it relates two effects, $\eta_1$ and $\eta_2$, such that
$\txn(\eta_1) = \txn(\eta_2)$ and $\id(\eta_1) < \id(\eta_2)$.  The
semantics assigns monotonically increasing identifiers to effects, as
defined by the $\id(\eta) > \maxId(\A)$ condition of
\rulelabel{E-Aux}.  Evaluation contexts ($\ectx_i$) for
transaction-bound terms are defined so as to enforce a deterministic
sequential order of execution within a transaction, leading to a
deterministic total order among effect ids, which defines the session
order relation. Visibility ($\visZ$) on the other hand relates effects
across concurrent transactions.  Intuitively, $\visZ$ relates $\eta_1$
to $\eta_2$ if and only if $\eta_1$ was \emph{visible} to the
operation that generated $\eta_2$ during its execution, thus effecting
its return value ($\rval(\eta_2)$). For example, a read operation over
\C{X} may pick the value ($\rval$) of the write effect with the highest id
among its visible effects (this is made possible by appropriately
defining $\interp{\cdot}$ in \rulelabel{E-Var}, as we show later).
Thus, the value of a read depends on what write effects it can
witness. An operation can only witness the effects of already
concluded operations, which varies between executions due to the
non-deterministic order of evaluating the parallel composition of
transactions.

A more notable source of non-determinism, however, is the
\rulelabel{E-Aux} rule, which allows the machine to expose an
arbitrary subset ($S$) of existing effects ($\A$) to the incoming
operation. In other words, the machine is not obligated to reveal the
effects of all previous operations to an incoming operation. This
relaxation allows the abstract machine to model the semantics of
weakly-consistent data stores. For instance, operations issued to an
eventually consistent ({\sc ec}) replicated store could be dispatched
to different replicas whose states may not be in any well-defined
relationship. By allowing operations to witness arbitrary subsets of
the global state, the semantics models the weak visibility properties
of such stores; we elaborate on the implication of this style of
definition in \S\ref{sec:store-consistency}.  Stronger visiblity
properties can be expressed by imposing well-formedness constraints
over $\visZ$ via the trace invariant ($\I$). Since the abstract
machine is obligated to satisfy $\I$ at every step of the execution,
operations are guaranteed to experience the level of isolation
specified by $\I$.  Thus, in executions that run to completion, the
abstract machine models a store that provides the required levels of
isolation.  Notably, the machine achieves this without defining an
operation semantics for isolation levels, solely by relying on their
declarative characterization as trace well-formedness constraints.
\S\ref{sec:ansi-isolation} shows the specification of various ANSI SQL
isolation levels stated as trace well-formedness constraints.

@@

As described previously, \rulelabel{E-Aux} abstracts away the
operation-specific behaviour of a machine step as a function ($\F$)
that accepts a set ($S$) of effects chosen by the machine to make
visible to the operation, interprets the operation w.r.t $S$, and
returns an appropriate effect that encodes its return value. Rules
\rulelabel{E-Var}, \rulelabel{E-Asgn} and \rulelabel{E-Commit} define
such functions for read, write and commit operations, respectively.
The effect returned by the function in each case includes its
transaction id ($T_i$) along with an arbitrarily chosen effect id
($j$) that is later verified to be unique in \rulelabel{E-Aux}. The
$\rval$ for a write is the value being written, and for commit it is
$\bot$. In case of a read, the value read depends on how the read
operation chooses to interpret the given set ($S$) of visible effects.
The interpretation may depend on application semantics. For example, a
monotonically increasing counter application may choose to let a write
with the largest value determine the value of a read. A more
straightforward interpretation has already been described before. To
accommodate multiple interpretations, the semantics is made parametric
over an interpretation function ($\interp{\cdot}$) that accepts a set
of effects and a variable name, and returns the value associated with
the variable. A straightforward interpretation function that chooses
the last write (i.e., write with largest id) is shown below:
